On Population-Based FA Analysis with Non-independent Normalization and Statistics

### 2011-06-08 ANTsTBSS/paper.tex
Minor edit.
ENH: adding the caveats paragraph

\title{Normalization Circularity in Population-Based FA Analysis}

@@

\begin{abstract}
Template-based normalization plays a critical role in many population
studies based on diffusion tensor imaging (DTI).  Several current
approaches compute the fractional anisotropy (FA) from the DTI and then
directly register the FA image to a FA template before performing
statistical comparison between different groups within a population.  In
this paper, we demonstrate that this analysis strategy induces selection
bias that varies in strength with the similarity metric used in
registration.  Specifically, we show that the popular sum of squared
difference intensity metric maximizes effect size, rather than
anatomical alignment, and consistently overestimates statistical
significance.  Consequently, we advocate a more traditional approach
that normalizes a population of FA images through the associated T1
anatomical images.  We show that the Symmetric Group Normalization
strategy for unbiased template construction from anatomical (e.g.
T1-weighted) images provides a valid and unbiased coordinate system for
statistical inference of FA differences and yet retains detection power
in comparison to FA-based normalization methods.  We establish this
outcome by using open data comprised of normal controls from the International
Neuroimaging Data-sharing Initiative as well as a cohort of 16 traumatic
brain injury survivors compared with 17 control subjects.  Alignment
comparisons with the normalization component of the popular tract-based
spatial statistics (TBSS) show that our anatomically-based template
approach produces good alignments despite the absence of any direct
FA-to-FA registration.  Of practical consideration we note that the
software tools discussed in this work are provided to the public as part
of the open source Advanced Normalization Tools (ANTs) repository.  In conclusion, we recommend against circular analysis strategies that select image similarity metrics that are tied directly to the ultimate statistical functions used to assess population effects.

@@

\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=85mm]{simulatedData.pdf}
\end{tabular}
\caption{(a) Axial slice from the discrete BrainWeb data set.  From (a) 
we create the (b) simulated template and sample (c) control and (d) patient images with both 
additive noise and subsequent Gaussian smoothing. }
\label{fig:simulated_data}
\end{figure}

\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=85mm]{simulatedDataWarped.pdf}
\end{tabular}
\caption{Warping the image in Fig. \ref{fig:simulated_data}(c) to the template in 
  Fig. \ref{fig:simulated_data}(b) using the SSD similarity metric produces the image shown in
  (a).  The corresponding deformation is given in (b).  Although the images were anatomically
  aligned prior to invoking ANTs registration, deformation is driven by the similarity metric
  causing the accumulation of Type 1 errors.}
\label{fig:simulated_data_warped}
\end{figure}

@@

\begin{figure*}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=180mm]{simulatedTesting.pdf}
\end{tabular}
\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the simulated experiment adjusted for multiple comparisons (FDR).
The true significance curves are drawn in black.  Elevated statistical power is seen with all 4 
         metrics for both tests although the effect is less in the more robust Wilcoxon results. Not shown are the significance curves for the Brunner-Munzel testing \cite{Rorden2007} which showed similar effects to the results on the left.
        }
\label{fig:simulated_testing}
\end{center}        
\end{figure*}

After creation of the simulated data set, we performed voxelwise testing (both $t$-testing
and Wilcoxon rank-sum testing, one-tailed, $\alpha = 0.95$) with false discovery
rate correction within the white matter to get ``true'' cumulative distribution functions (CDFs).  The CDF vertical axis shows the cumulative number of voxels with corrected $p$-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.  We denote these graphs as "statistical significance curves."  
The true curves for the simulated experiment are plotted in black in Fig. \ref{fig:simulated_testing}.  
We note that the amount of smoothing increases the proportion of significant voxels beyond
the prescribed 10\%.

@@

This example demonstrates the undesirable phenomenon previously described.  Even though these images were anatomically aligned prior to registration, the resulting deformations (falsely) increased detection power.  This effect can be explained in the context of Eqn. (\ref{eq:variance}) and by the fact that we added noise to the underlying distribution which allows the deformable mapping algorithms the freedom to transform the positions of the most "powerful" voxels to minimize variance across the population (and increase the $t$-test statistics).  Whereas  decreasing the standard deviation of the noise causes an increase in separation between  the SSD and Demon's metric pair with the MI and CC metric pair with the latter pair closer to the the true significance curve, increasing the noise yields the same ranking with a more even spread between significance curves.

@@

\begin{figure*}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=170mm]{workflow.pdf}
\end{tabular}
\caption{Possible ANTs-based TBSS workflow using SyGN.  In contrast to standard TBSS in which an individual subject's FA is registered to the FMRIB58 template
(transform represented with the blue arrow), our proposed modification first constructs an ANTs optimal anatomical template as illustrated in the center of the figure.  Each FA image is then rigidly mapped, with distortion correction, to the individual T1 using the white matter mask as shown on the left.  The individual T1-weighted image is then nonrigidly registered to the anatomical template.  Optionally, one can register the T1 template to the MNI152 template which resides in the same space as the FMRIB58 template.  Thus, the composite transform, carefully constructed without any FA-to-FA image registration, can take an individual subject's FA map (or other DTI-derived measure) to the desired standardized coordinate system.    }
\label{fig:workflow}
\end{center}
\end{figure*}

@@

\section{Results and Discussion}
%Results should be clear and concise.

@@

\begin{figure*}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=180mm]{tbiTesting.pdf}
\end{tabular}
\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).  Elevated statistical power is seen in the $t$-test with the similarity metrics with the most correspondence to Eqn. (\ref{eq:variance}), namely the SSD and Demons metric which shrinks drastically for the Wilcoxon rank-sum test.  Results for the other metrics including TBSS and the T1 template-based strategy (``antsT1'') remain fairly stable between the two tests.    TBSS and the 4 ANTs metric
results were all obtained using direct registration involving subject FA and the FMRIB58 template.  The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$ 
common T1 template $\rightarrow$ MNI152 (FMRIB58) template.    }
\label{fig:tbi_testing}
\end{center}        
\end{figure*}

@@

\begin{figure}
  \begin{center}
\begin{tabular}{c}
  \includegraphics[width=85mm]{tbiCoV.pdf}
\end{tabular}
\caption{Comparison of the different alignment strategies in normalizing the TBI data to the FMRIB58 template.  Assessment is quantitated using the voxelwise CV of the FA.  TBSS and the 4 ANTs metric
results were all obtained using direct registration involving subject FA and the FMRIB58 template.  The curve denoted as ``antsT1'' was also obtained in FMRIB58 space with the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$ 
common T1 template $\rightarrow$ MNI152 (FMRIB58) template. }
\label{fig:tbi_cov}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=85mm]{meanFA.pdf}
\end{tabular}
\caption{Qualitative comparison of mean FA images used in TBSS taken from the TBI data described in Section 2.  The mean FA image illustrated in (a) is created by mapping the sample FA population to the default FMRIB58 template which is the standard protocol for TBSS.  In comparison, the mean template represented in (c) is created from alignment of the population sample to the optimal T1 template as described in the paper. The respective skeleton masks in (b) and (d) are created by thresholding the resulting skeleton values $\geq 0.2$.  }
\label{fig:meanFA}
\end{center}
\end{figure}

@@

%\begin{figure}
%\begin{tabular}{c}
%  \includegraphics[width=80mm]{hoonData2.pdf}
%\end{tabular}
%\caption{Results of TBSS analysis using both the ANTs template-based approach (top row) and the standard TBSS protocol (bottom row).  Axial slices were chosen based on visual assessment of correspondence in both the ANTs template and MNI 152 template.
%On the left is the thresholded skeleton overlaid on the mean FA image.  
%On the right, significant areas of difference are given in red ($p \leq 0.01$ after the threshold-free cluster enhancement correction described in \cite{Smith2009}).  The yellow arrow points to an area of significant difference in the peripheral white matter found using ANTs-TBSS.
%}
%\label{fig:hoonData}
%\end{figure}

@@

 \begin{figure}
  \begin{center}
\begin{tabular}{c}
  \includegraphics[width=85mm]{NKI2.pdf}
\end{tabular}
\caption{Axial slice 80 of FA mean images used in the NKI alignment assessment.  (a) The standard FMRIB58 template. (b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).  
(c) and (d) The mean FA image produced using the ANTs template-based strategy employing 5 and 30 randomly chosen subjects, respectively, warped to the space of (a) followed by T1 registration of all subjects to the resulting anatomical template.  Note the visual alignment difference particularly in the peripheral white matter complex.
}
\label{fig:NKI}
\end{center}
\end{figure}

@@

%\begin{figure}
%\begin{tabular}{c}
%  \includegraphics[width=80mm]{NYU.pdf}
%\end{tabular}
%\caption{Axial slice 80 of FA mean images used in the NYU alignment assessment.  (a) The standard FMRIB58\_FA\_1mm template. (b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).  
%(c)-(f) The mean FA image produced using the ANTs template-based strategy employing 5, 10, 20, and 30 randomly chosen subjects, respectively, warped to the space of (a).  Note the visual alignment difference particularly in the peripheral white matter complex.
%}
%\label{fig:NYU}
%\end{figure}

@@

%\begin{figure*}
%\begin{tabular}{c}
%  \includegraphics[width=180mm]{NYUplots2.pdf}
%\end{tabular}
%\caption{{\bf NYU alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig. \ref{fig:NKI}(b)).  In addition to the whole brain analysis (first column), the CV was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).  The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
%}
%\label{fig:NYUplots}
%\end{figure*}

\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=85mm]{NKIplots3.pdf}
\end{tabular}
\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig. \ref{fig:NKI}(b)).  The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
}
\label{fig:NKIplots}
\end{figure}

@@

\begin{figure*}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=180mm]{nkiTesting.pdf}
\end{tabular}
\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results 
(one-tailed, $\alpha = 0.95$) for the NKI FA data aligned to the FMRIB58 template 
adjusted for multiple comparisons (FDR).  
Similar statistical power is seen in all alignment scenarios except for the more robust
metrics (CC and MI) in the $t$-test which show decreased power.   }
\label{fig:nki_testing}
\end{center}        
\end{figure*}

@@

\section{Discussion} Prior evaluation work using anatomically labeled
data indicates that mutual information,
cross-correlation, Demons, and SSD metrics are all capable of producing
high quality anatomical alignment (e.g. \cite{Klein2009}).  All of this work, though, compares
the quality of alignment at a relatively coarse scale of major gyri,
lobes and regions.  At the voxel level, it is nearly impossible to
determine the ground truth correspondence.  Indeed, there is no reason
to believe that subtle differences between a set of alignment strategies
(induced, in the experiments of this paper, by changing the similarity
metrics used in image registration) that all apparently work well should
be considered significant.  This work, however, highlights a dramatic impact of similarity metric choice on detection power in template-based FA studies.  Our contention is that this dramatic "improvement" in detection power is not due to better anatomical alignment.  Rather, it is a result of circularity/nonindependence of the normalization and statistical estimation strategy.  Consequently, we recommend that future work should take greater care when pairing the normalization strategy with the statistical framework.  Furthermore, we recommend that researchers should diligently seek to maintain independence between these two critical stages of a population study.  That is, {the features that drive the minimization of a similarity metric should be independent of the features that will be used in hypothesis testing.  In particular, one should avoid combining the SSD metric for normalizing FA with Student's $t$-test for FA differences.

>>>>

\title{On Population-Based FA Analysis with Non-independent Normalization and Statistics}

@@

\begin{abstract}
Template-based normalization plays a critical role in many population
studies based on diffusion tensor imaging (DTI).  Several current
approaches compute the fractional anisotropy (FA) from the DTI and then
directly register the FA image to a FA template before performing
statistical comparison between different groups within a population.  In
this paper, we demonstrate that this analysis strategy induces selection
bias that varies in strength with the similarity metric used in
registration.  Specifically, we show that the popular sum of squared
difference intensity metric maximizes effect size, rather than
anatomical alignment, and consistently overestimates statistical
significance.  Consequently, we advocate a more traditional approach
that normalizes a population of FA images through the associated 
anatomical images.  We show that the Symmetric Group Normalization
strategy for unbiased template construction from anatomical (e.g.
T1-weighted) images provides a valid and unbiased coordinate system for
statistical inference of FA differences and yet retains detection power
in comparison to FA-based normalization methods.  We establish this
outcome by using open data comprised of normal controls from the International
Neuroimaging Data-sharing Initiative as well as a cohort of 16 traumatic
brain injury survivors compared with 17 control subjects.  Alignment
comparisons with the normalization component of the popular tract-based
spatial statistics (TBSS) show that our anatomically-based template
approach produces good alignments despite the absence of any direct
FA-to-FA registration.  Of practical consideration we note that the
software tools discussed in this work are provided to the public as part
of the open source Advanced Normalization Tools (ANTs) repository.  In conclusion, we recommend against circular analysis strategies that select image similarity metrics that are tied directly to the ultimate statistical functions used to assess population effects.

@@

\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=85mm]{simulatedData.jpg}
\end{tabular}
\caption{(a) Axial slice from the discrete BrainWeb data set.  From (a) 
we create the (b) simulated template and sample (c) control and (d) patient images with both 
additive noise and subsequent Gaussian smoothing. }
\label{fig:simulated_data}
\end{figure}

\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=85mm]{simulatedDataWarped.jpg}
\end{tabular}
\caption{Warping the image in Fig. \ref{fig:simulated_data}(c) to the template in 
  Fig. \ref{fig:simulated_data}(b) using the SSD similarity metric produces the image shown in
  (a).  The corresponding deformation is given in (b).  Although the images were anatomically
  aligned prior to invoking ANTs registration, deformation is driven by the similarity metric
  causing the accumulation of Type 1 errors.}
\label{fig:simulated_data_warped}
\end{figure}

@@

\begin{figure*}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=180mm]{simulatedTesting.jpg}
\end{tabular}
\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the simulated experiment adjusted for multiple comparisons (FDR).
The true significance curves are drawn in black.  Elevated statistical power is seen with all 4 
         metrics for both tests although the effect is less in the more robust Wilcoxon results. Not shown are the significance curves for the Brunner-Munzel testing \cite{Rorden2007} which showed similar effects to the results on the left.
        }
\label{fig:simulated_testing}
\end{center}        
\end{figure*}

After creation of the simulated data set, we performed voxelwise testing (both $t$-testing
and Wilcoxon rank-sum testing, aka the Mann-Whitney test, one-tailed,
$\alpha = 0.95$) with false discovery rate correction within the white matter to get ``true'' cumulative distribution functions (CDFs).  The CDF vertical axis shows the cumulative number of voxels with corrected $p$-values at or below a given significance level (horizontal axis) and thus may be used as a surrogate measure for spatial sensitivity.  We denote these graphs as "statistical significance curves."  
The true curves for the simulated experiment are plotted in black in Fig. \ref{fig:simulated_testing}.  
We note that the amount of smoothing increases the proportion of significant voxels beyond
the prescribed 10\%.

@@

This example demonstrates the undesirable phenomenon previously
described.  Even though these images were anatomically aligned prior
to registration, the resulting deformations (falsely) increased
detection power.  This effect can be explained in the context of
Eqn. (\ref{eq:variance}) and by the fact that we added noise to the
underlying distribution which allows the deformable mapping algorithms
the freedom to transform the positions of the most "powerful" voxels
to minimize variance across the population (and increase the $t$-test
statistics).  Although we do not show the results, additional
experiments show that decreasing the standard deviation of the noise
causes an increase in separation between the SSD/Demon's metrics and
the MI and CC metrics.  MI and CC are closer to the the true
significance curve.  Alternatively, if we increase the noise level,
the same ranking occurs with even more spread between significance curves.

@@

\begin{figure*}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=170mm]{workflow.jpg}
\end{tabular}
\caption{Possible ANTs-based TBSS workflow using SyGN.  In contrast to standard TBSS in which an individual subject's FA is registered to the FMRIB58 template
(transform represented with the blue arrow), our proposed modification first constructs an ANTs optimal anatomical template as illustrated in the center of the figure.  Each FA image is then rigidly mapped, with distortion correction, to the individual T1 using the white matter mask as shown on the left.  The individual T1-weighted image is then nonrigidly registered to the anatomical template.  Optionally, one can register the T1 template to the MNI152 template which resides in the same space as the FMRIB58 template.  Thus, the composite transform, carefully constructed without any FA-to-FA image registration, can take an individual subject's FA map (or other DTI-derived measure) to the desired standardized coordinate system.    }
\label{fig:workflow}
\end{center}
\end{figure*}

@@

\section{Results}
%Results should be clear and concise.

@@

\begin{figure*}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=180mm]{tbiTesting.jpg}
\end{tabular}
\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results (one-tailed, $\alpha = 0.95$) for the TBI FA data aligned to the FMRIB58 template adjusted for multiple comparisons (FDR).  Elevated statistical power is seen in the $t$-test with the similarity metrics with the most correspondence to Eqn. (\ref{eq:variance}), namely the SSD and Demons metric which shrinks drastically for the Wilcoxon rank-sum test.  Results for the other metrics including TBSS and the T1 template-based strategy (``antsT1'') remain fairly stable between the two tests.    TBSS and the 4 ANTs metric
results were all obtained using direct registration involving subject FA and the FMRIB58 template.  The antsT1 results were also obtained in FMRIB58 space but the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$ 
common T1 template $\rightarrow$ MNI152 (FMRIB58) template.    }
\label{fig:tbi_testing}
\end{center}        
\end{figure*}

@@

\begin{figure}
  \begin{center}
\begin{tabular}{c}
  \includegraphics[width=85mm]{tbiCoV.jpg}
\end{tabular}
\caption{Comparison of the different alignment strategies in normalizing the TBI data to the FMRIB58 template.  Assessment is quantitated using the voxelwise CV of the FA.  TBSS and the 4 ANTs metric
results were all obtained using direct registration involving subject FA and the FMRIB58 template.  The curve denoted as ``antsT1'' was also obtained in FMRIB58 space with the following transformation composition: subject FA $\rightarrow$ subject T1 $\rightarrow$ group T1 template $\rightarrow$ 
common T1 template $\rightarrow$ MNI152 (FMRIB58) template. }
\label{fig:tbi_cov}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=85mm]{meanFA.jpg}
\end{tabular}
\caption{Qualitative comparison of mean FA images used in TBSS taken from the TBI data described in Section 2.  The mean FA image illustrated in (a) is created by mapping the sample FA population to the default FMRIB58 template which is the standard protocol for TBSS.  In comparison, the mean template represented in (c) is created from alignment of the population sample to the optimal T1 template as described in the paper. The respective skeleton masks in (b) and (d) are created by thresholding the resulting skeleton values $\geq 0.2$.  }
\label{fig:meanFA}
\end{center}
\end{figure}

@@

%\begin{figure}
%\begin{tabular}{c}
%  \includegraphics[width=80mm]{hoonData2.jpg}
%\end{tabular}
%\caption{Results of TBSS analysis using both the ANTs template-based approach (top row) and the standard TBSS protocol (bottom row).  Axial slices were chosen based on visual assessment of correspondence in both the ANTs template and MNI 152 template.
%On the left is the thresholded skeleton overlaid on the mean FA image.  
%On the right, significant areas of difference are given in red ($p \leq 0.01$ after the threshold-free cluster enhancement correction described in \cite{Smith2009}).  The yellow arrow points to an area of significant difference in the peripheral white matter found using ANTs-TBSS.
%}
%\label{fig:hoonData}
%\end{figure}

@@

 \begin{figure}
  \begin{center}
\begin{tabular}{c}
  \includegraphics[width=85mm]{NKI2.jpg}
\end{tabular}
\caption{Axial slice 80 of FA mean images used in the NKI alignment assessment.  (a) The standard FMRIB58 template. (b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).  
(c) and (d) The mean FA image produced using the ANTs template-based strategy employing 5 and 30 randomly chosen subjects, respectively, warped to the space of (a) followed by T1 registration of all subjects to the resulting anatomical template.  Note the visual alignment difference particularly in the peripheral white matter complex.
}
\label{fig:NKI}
\end{center}
\end{figure}

@@

%\begin{figure}
%\begin{tabular}{c}
%  \includegraphics[width=80mm]{NYU.jpg}
%\end{tabular}
%\caption{Axial slice 80 of FA mean images used in the NYU alignment assessment.  (a) The standard FMRIB58\_FA\_1mm template. (b) The mean FA image of all aligned 57 NKI data using the TBSS default FLIRT/FNIRT parameters optimized for registration to (a).  
%(c)-(f) The mean FA image produced using the ANTs template-based strategy employing 5, 10, 20, and 30 randomly chosen subjects, respectively, warped to the space of (a).  Note the visual alignment difference particularly in the peripheral white matter complex.
%}
%\label{fig:NYU}
%\end{figure}

@@

%\begin{figure*}
%\begin{tabular}{c}
%  \includegraphics[width=180mm]{NYUplots2.jpg}
%\end{tabular}
%\caption{{\bf NYU alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig. \ref{fig:NKI}(b)).  In addition to the whole brain analysis (first column), the CV was separately calculated over the inner and outer portions of the white matter (second and third columns, respectively).  The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
%}
%\label{fig:NYUplots}
%\end{figure*}

\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=85mm]{NKIplots3.jpg}
\end{tabular}
\caption{{\bf NKI alignment assessment:}  To compare alignments between the  standard TBSS approach of using the FMRIB58 template versus using the different ANTs templates, we calculated the CV at each voxel (for both FA and MD) and plotted the CDF over the range of values $[0,0.8]$ accumulated over all the voxels inside the thresholded (FA $\geq 0.2$) standard TBSS mean FA image (see Fig. \ref{fig:NKI}(b)).  The CDF plots for the ANTs templates are all very similar (although trending towards slight improvement with larger template sample size) and illustrate the general improvement in FA and MD alignment over registration of each FA image to the FMRIB58 template.
}
\label{fig:NKIplots}
\end{figure}

@@

\begin{figure*}
\begin{center}
\begin{tabular}{c}
  \includegraphics[width=180mm]{nkiTesting.jpg}
\end{tabular}
\caption{Student's $t$-test and the nonparametric Wilcoxon rank-sum test results 
(one-tailed, $\alpha = 0.95$) for the NKI FA data aligned to the FMRIB58 template 
adjusted for multiple comparisons (FDR).  
Similar statistical power is seen in all alignment scenarios except for the more robust
metrics (CC and MI) in the $t$-test which show decreased power.   }
\label{fig:nki_testing}
\end{center}        
\end{figure*}

@@

\section{Discussion} Prior evaluation work using anatomically labeled
data indicates that mutual information, cross-correlation, Demons, and
SSD metrics are all capable of producing high quality anatomical
alignment (e.g. \cite{Klein2009}).  All of this work, though, compares
the quality of alignment at a relatively coarse scale of major gyri,
lobes and regions.  At the voxel level, it is nearly impossible to
determine the ground truth correspondence.  Indeed, there is no reason
to believe that subtle differences between a set of alignment
strategies (induced, in the experiments of this paper, by changing the
similarity metrics used in image registration) that all apparently
work well should be considered significant.  This work, however,
highlights a dramatic impact of similarity metric choice on detection
power in template-based FA studies.  Our contention is that this
dramatic "improvement" in detection power is not due to better
anatomical alignment.  Rather, it is a result of
circularity/nonindependence of the normalization and statistical
estimation strategy.  Consequently, we recommend that future work
should take greater care when pairing the normalization strategy with
the statistical framework.  Furthermore, we recommend that researchers
should diligently seek to maintain independence between these two
critical stages of a population study.  That is, the features that
drive the minimization of a similarity metric should be independent of
the features that will be used in hypothesis testing.  In particular,
one should avoid combining the SSD metric for normalizing FA with
Student's $t$-test for FA differences.  A final point is that
cumulative distribution functions and other power assessment methods
cannot be regarded as meaningful comparison techniques in the presence
of normalization/statistical non-independence.  Similar issues are
raised in \cite{Kriegeskorte2010}.

Caveats with our analysis include a lack of ground truth in the
clinical component of the study.   Additionally, we are not addressing
our results' biological plausibility, nor do we
intend to (in this paper) enter into discussion on the mechanisms of
TBI, gender difference and DTI that may lead to the detected results.
Rather, we focus on the technical issue of circularity induced by
normalizing the image feature of interest with functions that maximize
the statistic used in hypothesis testing.  We also cannot argue that our analysis
invalidates previous results that used a circular strategy.  Instead, we encourage the community to
reexamine FA normalization and analysis methods and to consider the
potential confounds highlighted in this paper.  Rank tests
provide a more conservative alternative as does the methodology
presented here based on T1 to DT intrasubject mapping and T1
normalization.
